---
title: 'Statistical Inference: Simulating Exponential Distributions'
author: "Henrique Souza (github.com/htssouza)"
output: pdf_document
---

# Overview

This report is intended to investigate the exponential distribution in R
and compare it with the Central Limit Theorem (CLT).

# Simulations

For all the simulations lambda will be equal to 0.2.
A seed will also be set.

```{r echo = TRUE}
lambda <- 0.2
set.seed(123456)
```

Now lets generate some sample values and plot their histogram.

```{r echo = TRUE}
sample <- rexp(40, lambda)
hist(sample,
     main = "Histogram of a sample exponential distribution",
     xlab = "Sample values",
     col = c("red"),
     breaks = 10)
```

# Sample Mean versus Theoretical Mean

```{r echo = TRUE}
sample_mean <- mean(sample)
theoretical_mean <- 1 / lambda
```

You can see that their values is pretty close.

Let's test with a larger sample size (1000) and see if it gets even closer:

```{r echo = TRUE}
larger_sample <- rexp(1000, lambda)
larger_sample_mean <- mean(larger_sample)
```

It is closer to the theoretical value.

# Sample Variance versus Theoretical Variance

Let's compute the sd of our sample.

```{r echo = TRUE}
sample_sd <- sd(sample)
theoretical_sd <- 1 / lambda
```

Similar but quite far from the expected value.

Let's now repeat the comparison with the larger sample.

```{r echo = TRUE}
larger_sample_sd <- sd(larger_sample)
```

Which is much better.

# Distribution

Computing and ploting the histogram of a thousand exponential distributions.

```{r echo = TRUE}
means = NULL
for (i in 1 : 1000) means = c(means, mean(rexp(40, lambda)))
hist(means,
     main = "Histogram of mean of samples of exponential distributions",
     xlab = "Means",
     col = c("red"),
     breaks = 20)
```

Plotting that with a normal distribution with the same mean and standard deviation.

```{r echo = TRUE}
ndist <- rnorm(1000, mean = mean(means), sd = sd(means))

data <- data.frame(value = means)
data$type <- "expotential"
data <- rbind(data, data.frame(value = ndist, type = rep("normal", length(ndist))))

library(ggplot2)
ggplot(data, aes(x=value, fill=type)) +
    geom_histogram(
      binwidth=.5,
      alpha=.5,
      position="identity",
      breaks=seq(
        min(data$value),
        max(data$value),
        by=((max(data$value) - min(data$value))/20)
        )
      )
```
