---
title: 'Statistical Inference: Tooth Growth Analysis'
author: "Henrique Souza (github.com/htssouza)"
output:
  pdf_document:
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
  html_document: default
---

# Goals

1-) Load the ToothGrowth data and perform some basic exploratory data analyses 
2-) Provide a basic summary of the data
3-) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4-) State your conclusions and the assumptions needed for your conclusions

# 1-) Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r echo = TRUE}
# Load data
data(ToothGrowth)

# Data summary
str(ToothGrowth)

# A small sample of the data
head(ToothGrowth)
```

# 2-) Provide a basic summary of the data.

Dose is a factor (few different valyes) so make the preprocessing conversion:

```{r echo = TRUE}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

Current data as a box-plot:

```{r echo = TRUE}
library("lattice")
bwplot(len ~ dose | supp,data=ToothGrowth)
```

# 3-) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.



