---
title: 'Statistical Inference: Tooth Growth Analysis'
author: "Henrique Souza (github.com/htssouza)"
output:
  html_document: default
  pdf_document:
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
---

# Goals

1-) Load the ToothGrowth data and perform some basic exploratory data analyses 
2-) Provide a basic summary of the data
3-) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4-) State your conclusions and the assumptions needed for your conclusions

# 1-) Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r echo = TRUE}
# Load data
data(ToothGrowth)

# Data summary
str(ToothGrowth)

# A small sample of the data
head(ToothGrowth)
```

# 2-) Provide a basic summary of the data.

Comparing supplement results by dose:

```{r echo = FALSE}
library(ggplot2)
data <- data.frame(len = ToothGrowth$len, supp = ToothGrowth$supp, dose = as.factor(ToothGrowth$dose))
g <- ggplot(data, aes(x = dose, y = len, group = dose))
g <- g + labs(title = "Comparing supplements", x = "dose", y = "tooth length")
g <- g + geom_boxplot(aes(fill = dose))
g <- g+ facet_grid(. ~ supp)
g
```

# 3-) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

Summarizing data to compute confidence intervals (using Student T interval):

```{r echo = TRUE}
library(plyr)
summaries <- ddply(data, .(supp, dose), summarize, mean = mean(len), sd = sd(len), n = length(len))
summaries
```

Computing confidence intervals (95% of confidence):

```{r echo = TRUE}
summaries$min <- summaries$mean - (qt(.975, (summaries$n-1) * summaries$sd / sqrt(summaries$n)))
summaries$max <- summaries$mean + (qt(.975, (summaries$n-1) * summaries$sd / sqrt(summaries$n)))
summaries
```

Plotting these intervals:

```{r echo = TRUE}
g <- ggplot(summaries)
g <- g + labs(title = "Comparing length (confidence intervals) by dose", x = "supplement", y = "tooth length")
g <- g + geom_crossbar(
  aes(ymin = min, ymax = max, x = supp, y = min, fill = dose),
  fatten = 0
  )
g <- g + facet_wrap(~dose)
g
```

# 4-) Conclusions

Based on the summary above we can say (with 95% of sure that):

- lengths when using OJ (.5 dose) > VC (.5 dose)
- lengths when using OJ (1.0 dose) > VC (1.0 dose)
- lengths when using 2.0 of both OJ and VC are similar/equivalent

The assumptions are:

- current dataset (small, with only 60 rows)
- 95% of confidence


