---
title: 'Statistical Inference: Tooth Growth Analysis'
author: "Henrique Souza (github.com/htssouza)"
output:
  pdf_document:
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
  html_document: default
---

# Goals

1-) Load the ToothGrowth data and perform some basic exploratory data analyses 
2-) Provide a basic summary of the data
3-) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4-) State your conclusions and the assumptions needed for your conclusions

# 1-) Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r echo = TRUE}
# Load data
data(ToothGrowth)

# Data summary
str(ToothGrowth)

# A small sample of the data
head(ToothGrowth)
```

# 2-) Provide a basic summary of the data.

Dose is a factor (few different valyes) so make the preprocessing conversion:

```{r echo = TRUE}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

Plotting some histograms:

```{r echo = TRUE}
library(plyr)
library(ggplot2)
qplot(
  len,
  data = ToothGrowth,
  facets = .~supp~dose,
  binwidth = 1,
  geom="histogram")
```

# 3-) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

Summarizing data to compute confidence intervals (using Student T interval):

```{r echo = TRUE}
summaries <- ddply(ToothGrowth, .(supp, dose), summarize, mean = mean(len), sd = sd(len), n = length(len))
summaries
```

Computing confidence intervals (95% of confidence):

```{r echo = TRUE}
summaries$min <- summaries$mean - (qt(.975, (summaries$n-1) * summaries$sd / sqrt(summaries$n)))
summaries$max <- summaries$mean + (qt(.975, (summaries$n-1) * summaries$sd / sqrt(summaries$n)))
summaries
```

Plotting these intervals:

```{r echo = TRUE}
gg <- ggplot(summaries)
gg <- gg + geom_crossbar(
  aes(ymin = min, ymax = max, x = supp, y = min),
  fill = "blue",
  fatten = 0
  )
gg <- gg + facet_wrap(~dose)
gg
```

# 4-) Conclusions

Based on the summary above we can say (with 95% of sure that):

- lengths when using OJ (.5 dose) > VC (.5 dose)
- lengths when using OJ (1.0 dose) > VC (1.0 dose)
- lengths when using 2.0 of both OJ and VC are similar/equivalent

The assumptions are:

- current dataset (small, with only 60 rows)
- 95% of confidence


